---
title: "64036_Assignment1"
author: "Durgaprasad Gandi"
date: "2023-09-29"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library("knitr")
```
# Summary of the Carseats dataset:
```{r}
library(ISLR) #loads the ISLR library
summary(Carseats)
```
# Number of (observations)rows in Carseats dataset:
```{r}
number_rows <- nrow(Carseats) #To get number of rows in the dataset.
print(number_rows)
```
***We get 400 observations in a Carseats dataset***

# Maximum value of the advertising attribute:

```{r}
Advertising_Max <-max(Carseats$Advertising) #max is used to get maximum value of dataset.
print(Advertising_Max)
```
# Calculating the inter quartile of price attribute:
```{r}
iqr_price <-IQR(Carseats$Price) #IQR is function to get inter quartile range.
print(iqr_price)
```
***The value of IQR is 31***

# Plotting of Price against Sales:
```{r}
plot_SP<- plot(Carseats$Price,
Carseats$Sales,
xlab="Price", # on x axis we take Price as label.
ylab="Sales") #on y axis we take Sales as label.
print(plot_SP)
```
# Correlation of Price and Sales:
```{r}
cor_cooeft <- cor(Carseats$Price, Carseats$Sales) #$ indicates respective column of dataset.
print(cor_cooeft)
```
***As the value is -0.4449 i.e the relation between Price and Sales is negative which indicates Weak relation***


